%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cr                                                                       
%cr            (C) Copyright 1995 The Board of Trustees of the            
%cr                        University of Illinois                         
%cr                         All Rights Reserved                           
%cr                                                                       
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RCS INFORMATION:
%
%       $RCSfile: ig_intro.tex,v $
%       $Author: johns $        $Locker:  $                $State: Exp $
%       $Revision: 1.9 $      $Date: 2000/05/16 18:52:35 $
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DESCRIPTION:
%
% INSTALLATION GUIDE : introduction chapter
%   1) General description of document and purpose
%   2) Organization of document
%   3) How to use this guide
%   4) For more information ...
%   5) Contacting the authors
%   6) Referencing this work
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\label{chapter:ig:intro}

% general intro text
\VMD\ is a molecular visualization and analysis program designed to
be used for interactive display of molecular systems, particularly
biopolymers such as proteins, nucleic acids, and biological
assemblies such as membrane lipid bilayers.  As described in the
\VMD\ Programmer's and User's Guides, \VMD\ has several goals:
\begin{itemize}
  \item General molecular visualization;
  \item Visualization of dynamic molecular data;
  \item Analysis of molecular structures;
  \item Display and control of molecular dynamics simulations;
  \item Support for several input and display (output) devices,
       including output image files;
  \item Modifiable and extensible program structure.
\end{itemize}

This manual, along with the Users
Guide and Programmers Guide, documents the use of \VMD.

% how the guide is organized
\section{Guide organization}

This \DOCTITLE\ details how to compile and install \VMD.  You have the
option of simply installing a precompiled version of
\VMD\ provided in the standard distribution, or the option of recompiling \VMD\
with a different set of features and options.  Depending on the method
preferred, different chapters of this Guide will be relevant.  This
manual is organized into the following chapters following this
introduction:

\begin{itemize}
  \BOLDLISTITEM{Obtaining and unpacking \VMD\ (chapter
\ref{chapter:ig:unpack})} The first step in installing \VMD;
this chapter should be read first.

  \BOLDLISTITEM{Compiling \VMD\ (chapter \ref{chapter:ig:compile})}
Describes how to compile \VMD.  The program may be compiled for a
variety of operating system versions, and may optionally include
or exclude a number of features.  Chapter \ref{chapter:ig:compile}
lists all the operating systems for which \VMD\ may be compiled,
describes all the optional program features which may be included
in the compiled executable, and mentions all the configurable
parameters which may be preferentially customized.  The
\VMD\ executables provided with the standard distribution are compiled
with a default set of features, as described in chapter
\ref{chapter:ig:install}.

  \BOLDLISTITEM{Installing \VMD\ (chapter \ref{chapter:ig:install})} 
Describes how to install the \VMD\ executables and data files, once
\VMD\ has been properly compiled.  The standard distribution includes
precompiled versions of \VMD\ for several different Unix operating
systems, as listed in table \ref{table:ig:execs}.  You may install
one of these precompiled versions as is, or you may refer to
chapter \ref{chapter:ig:compile} to first compile a new version of \VMD\
before installing it.

  \BOLDLISTITEM{Customizing \VMD\ (chapter \ref{chapter:ig:custom})} 
Mentions how to edit the \VMD\ configuration data files in order to
customize the appearance and behavior of \VMD\ for a particular
environment.  This is done after \VMD\ has been successfully compiled
and installed.
\end{itemize}

\section{How to use this guide}
\label{section:ig:howtouse}

There are two ways to install \VMD:
\begin{enumerate}
  \item Use a provided precompiled version of \VMD; to do this, use
chapter \ref{chapter:ig:unpack} and chapter \ref{chapter:ig:install}. 
Currently, precompiled versions of \VMD\ are supplied for the following
operating systems (see table \ref{table:ig:execs}):
  \begin{itemize}
    \item SGI IRIX 5.X or IRIX 6.X.
  \end{itemize}
For other operating systems, the second method must be used. 
  \item Compile \VMD\ with a selected set of features and
parameters.  This requires first unpacking the distribution (chapter
\ref{chapter:ig:unpack}), compiling the program (chapter
\ref{chapter:ig:compile}), and then installing the executables and
data files (chapter \ref{chapter:ig:install}).  Chapter
\ref{chapter:ig:compile} lists the operating systems for which
\VMD\ is supported, and all possible program features and
user-selectable parameters.  You are also strongly advised to
get and install the FLTK or XForms, Tcl, and Tcl-X packages.
\end{enumerate}

Finally, after installation is complete, several factors may be customized
about how \VMD\ appears to users upon startup.
This is done by following the instructions in chapter \ref{chapter:ig:custom}.

% for more information ...
\input{vmd_moreinfo}
\label{section:ig:moreinfo}

% contacting the authors
\input{vmd_authors}
\label{section:ig:authors}

% credits, and referencing this work
\input{vmd_ref}
\label{section:ig:ref}

