%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cr                                                                       
%cr            (C) Copyright 1995 The Board of Trustees of the            
%cr                        University of Illinois                         
%cr                         All Rights Reserved                           
%cr                                                                       
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RCS INFORMATION:
%
%       $RCSfile: ug_intro.tex,v $
%       $Author: johns $        $Locker:  $                $State: Exp $
%       $Revision: 1.32 $      $Date: 2014/12/28 22:57:43 $
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DESCRIPTION:
%
% USERS GUIDE : introduction chapter
%   1) General description of document and purpose
%   2) Organization of document
%   4) For more information ...
%   5) Contacting the authors
%   6) Referencing this work
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}

% general intro text
\VMD~\cite{HUMP96} is a molecular graphics program designed for the
interactive visualization and analysis of biopolymers such as proteins,
nucleic acids, lipids, and membranes.  \VMD\ runs on all major 
Unix workstations, Apple MacOS X, and Microsoft Windows.
Online information about VMD is available from:
\\ 
\htmladdnormallink{{\tt http://www.ks.uiuc.edu/Research/vmd/}}
		   {http://www.ks.uiuc.edu/Research/vmd/}
\vspace{0.1in}
\\
List of key \VMD\ features:
\begin{itemize}
  \BOLDLISTITEM{General molecular visualization} At its heart, \VMD\
is a general application for displaying molecules containing
any number of atoms and is similar to other molecular visualization 
programs in its basic capabilities.
\VMD\ reads data files using an extensible plugin system, and supports
Babel\mycite{}{} for conversion of other formats.
User-defined atom selections can be displayed in any of the
standard molecular representations.
Displayed graphics can be exported to an image file, 
to a scene file usable by ray tracing programs, or to
a geometry description file suitable for use with 3-D printers.

\BOLDLISTITEM{Visualization of dynamic molecular data}
\VMD\ can load atomic coordinate trajectories from 
AMBER, Charmm, DLPOLY, Gromacs, MMTK, NAMD, X-PLOR, and many other 
simulation packages.  The data can be used to animate the molecule or 
to plot the change in molecular properties such as angles, dihedrals, 
interatomic distances, or energies over time.

\BOLDLISTITEM{Visualization of volumetric data}
\VMD\ can load, generate, and display, volumetric maps.  Supported
map formats include CryoEM maps, electrostatic potential maps, 
electron density maps, and many other map file formats.

\BOLDLISTITEM{Interactive molecular dynamics simulations} 
\VMD\ can be used as a graphical front-end to a live molecular
dynamics program running on a remote supercomputer or
high-performance workstation.  \VMD\ can interactively apply
and visualize forces in an MD simulation as it runs.

\BOLDLISTITEM{Molecular analysis commands}
Many commands are provided for molecular analysis.  
These include commands to extract information on sets of atoms and molecules, 
vector and matrix routines for coordinate manipulation, and 
functions for computing values such as center of mass and radius 
of gyration.

\BOLDLISTITEM{Tcl and Python scripting languages}
\VMD\ uses the freely available Python and Tcl scripting languages 
for processing text commands.  These popular languages which contain
variables, loops, subroutines, and much more.
\VMD\ also uses the Tk Toolkit - a simple user interface toolkit 
that interfaces with Tcl.

\BOLDLISTITEM{Easy to extend}
\VMD\ is written in C and C++ and employs object-oriented design.
\VMD\ implements a plugin interface for extending its file format
support and for general purpose extensions in functionality.

\BOLDLISTITEM{Support for multimodal input and various display systems}
A number of different visual display and control systems are
supported in addition to the usual monitor, keyboard, and mouse.  
The VRPN library\mycite{}{} is used to get position and orientation
information from a wide variety of spatial input devices, including 
magnetic trackers, haptic (force feedback) devices, Spaceballs, etc.
\VMD\ works with WireGL and Chromium on tiled display walls,
and immersive VR environments via compiled-in CAVE and FreeVR support.


\end{itemize}

% contacting the authors
\input{vmd_authors}

% credits, and referencing this work
\input{vmd_ref}

% for more information ...
\input{vmd_moreinfo}


