%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cr                                                                       
%cr            (C) Copyright 1995-2009 The Board of Trustees of the            
%cr                        University of Illinois                         
%cr                         All Rights Reserved                           
%cr                                                                       
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RCS INFORMATION:
%
%       $RCSfile: vmd_ref.tex,v $
%       $Author: johns $        $Locker:  $                $State: Exp $
%       $Revision: 1.74 $      $Date: 2014/12/29 03:25:57 $
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DESCRIPTION:
%
% referencing this work, and credits and acknowledgments
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Registering \VMD}

\VMD\ is made available free of charge for all interested 
end-users of the software (but please see the Copyright and 
Disclaimer notices).  Please check the current \VMD\ license 
agreement for details.  Registration is part of our software download
procedure.  Once you've filled out the forms on the \VMD\ download
area and have read and agreed to the license, you are finished with
the registration process.

\section{Citation Reference}
The authors request that any published work or images created using \VMD\ 
include the following reference:

\begin{description}
  \item{Humphrey, W., Dalke, A. and Schulten, K.,} ``VMD - Visual Molecular
Dynamics'' {\em J. Molec. Graphics} {\bf 1996}, {\em 14.1}, 33-38.
\end{description}

\VMD\ has been developed by the 
Theoretical and Computational Biophysics Group at the
Beckman Institute for Advanced Science and Technology of the
University of Illinois at Urbana-Champaign.
This work is supported by the National Institutes of Health under
grant numbers NIH~9P41GM104601 and 5R01GM098243-02.

\section{Acknowledgments}

The authors would particularly like to thank those individuals 
who have contributed suggestions and improvements, particularly 
those contributing new features.  Special thanks go to 
Joshua Anderson, Anton Arkhipov, Andrew Dalke, Michael Bach, 
Alexander Balaeff, Ilya Balabin, Robert Brunner,
Eamon Caddigan, Jordi Cohen, Simon Cross,
Markus Dittrich, John Eargle, 
Peter Freddolino, Todd Furlong, Luis Gracia, 
Paul Grayson, Justin Gullingsrud, James Gumbart,
David Hardy, Konrad Hinsen,
Barry Isralewitz, Sergei Izrailev, Robert Johnson,
Axel Kohlmeyer, Michael Krone, Michelle Kuttell, Benjamin Levine,
John Mongan, Jim Phillips, Elijah Roberts, Jan Saam, 
Charles Schwieters, Marcos Sotomayor, Alexander Spaar,
John E. Stone, Johan Strumpfer,
Alexey Titov, Leonardo Trabuco, Dan Wright, and Kirby Vandivort.
The entire \VMD\ user community now benefits from your contributions.

The authors would like to thank individuals
who have indirectly helped with development 
by making suggestions, pushing for new features, and trying out buggy code.  
Thanks go to 
Aleksei Aksimentiev, Daniel Barsky, Axel Berg, 
Tom Bishop, Robert Brunner, Ivo Hofacker, Mu Gao, 
James Gumbart, Xiche Hu, Tim Isgro, Dorina Kosztin, Ioan Kosztin, 
Joe Landman, Ilya Logunov, Clare Macrae, Amy Shih, 
Lukasz Salwinski, Stephen Searle, Charles Schwieters, 
Ari Shinozaki, Svilen Tzonev, Emad Tajkhorshid, Michael Tiemann, 
Elizabeth Villa, Raymond de Vries, Simon Warfield,
Willy Wriggers, Dong Xu, and Feng Zhou.  

Many external libraries and packages are used in \VMD, and the 
program would not be as capable without them.  
The authors of \VMD wish to thank
the authors of FLTK;
the authors of Tcl and Tk; 
the authors of Python; 
the authors of VRPN;
Jon Leech for uniform point distributions;
Amitabh Varshney for SURF;
Dmitrij Frishman for developing STRIDE; 
Jack Lund for the url\_get perl script; 
Brad Grantham for the ACTC triangle consolidation library;
John E. Stone for the Tachyon ray tracer, WorkForce threading and timer routines, hash table code, and Spaceball drivers;
\index{rendering!Tachyon}
\index{spaceball!driver}
and 
Ethan Merrit for one of the ribbon drawing algorithms. 


\section{Copyright and Disclaimer Notices}
\index{VMD!copyright}
\index{copyright}
\begin{center}
VMD is  Copyright \copyright\ 1995-2015 
Theoretical and Computational Biophysics Group and the \\
Board of Trustees of the University of Illinois
\end{center}

\noindent
Portions of this code are copyright \copyright\ 1997-1998 Andrew Dalke.

\bigskip

The terms for using, copying, modifying, and distributing VMD are
specified by the VMD License.  The license agreement is distributed
with VMD in the file LICENSE. If for any reason you do not have this
file in your distribution, it can be downloaded from: \\ 
{\tt http://www.ks.uiuc.edu/Research/vmd/current/LICENSE.html} 
\\
Some of the code and executables used by \VMD\ have their own 
usage restrictions:  
\begin{itemize}

\LISTITEM{ACTC}\index{ACTC}
ACTC, the triangle consolidation library used in some versions of VMD,
is Copyright (C) 2000, Brad Grantham and Applied Conjecture,
all rights reserved.

Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions
are met:
\\
  1. Redistributions of source code must retain the above copyright
     notice, this list of conditions and the following disclaimer.
\\
  2. Redistributions in binary form must reproduce the above copyright
     notice, this list of conditions and the following disclaimer in the
     documentation and/or other materials provided with the distribution.
\\
  3. All advertising materials mentioning features or use of this software
     must display the following acknowledgment:
       This product includes software developed by Brad Grantham and
       Applied Conjecture.
\\
  4. Neither the name Brad Grantham nor Applied Conjecture
     may be used to endorse or promote products derived from this software
     without specific prior written permission.
\\
  5. Notification must be made to Brad Grantham about inclusion of this
     software in a product including the author of the product and the name
     and purpose of the product.  Notification can be made using email
     to Brad Grantham's current address (grantham@plunk.org as of September
     20th, 2000) or current U.S. mail address.

\LISTITEM{Python}\index{Python}
Python is made available subject to the terms and conditions in CNRI's
License Agreement. This Agreement together with Python may be
obtained from a proxy server on the Internet using the following
URL: {\tt http://hdl.handle.net/1895.22/1012}

\LISTITEM{PCRE}\index{PCRE}
The Perl Compatible Regular Expressions (PCRE) library used in VMD
was written by Philip Hazel and is Copyright (c) 1997-1999 
University of Cambridge.  
\\
Permission is granted to anyone to use this software for any purpose on any
computer system, and to redistribute it freely, subject to the following
restrictions:
\\
1. This software is distributed in the hope that it will be useful,
   but WITHOUT ANY WARRANTY; without even the implied warranty of
   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
\\
2. The origin of this software must not be misrepresented, either by
   explicit claim or by omission.
\\
3. Altered versions must be plainly marked as such, and must not be
   misrepresented as being the original software.
\\
4. If PCRE is embedded in any software that is released under the GNU
   General Purpose License (GPL), then the terms of that license shall
   supersede any condition above with which it is incompatible.

\LISTITEM{STRIDE}\index{stride}
STRIDE, the program used for secondary structure calculation, is
free to both academic and commercial sites provided that STRIDE will
not be a part of a package sold for money.  The use of STRIDE in
commercial packages is not allowed without a prior written
commercial license agreement.  See
{\tt http://www.embl-heidelberg.de/argos/stride/stride\_info.html}

\LISTITEM{SURF}\index{surf}
The source code for SURF is copyrighted by the original author,
Amitabh Varshney, and the University of North Carolina at Chapel Hill.
Permission to use, copy, modify, and distribute this software and its
documentation for educational, research, and non-profit purposes is
hereby granted, provided this notice, all the source files, and the
name(s) of the original author(s) appear in all such copies.
\\
BECAUSE THE CODE IS PROVIDED FREE OF CHARGE, IT IS PROVIDED "AS IS" AND
WITHOUT WARRANTY OF ANY KIND, EITHER EXPRESSED OR IMPLIED.
\\
This software was developed and is made available for public use with
the support of the National Institutes of Health, National Center for
Research Resources under grant RR02170.

\LISTITEM{Tachyon}\index{Tachyon}
The Tachyon multiprocessor ray tracing system and derivative code built
into VMD is Copyright (c) 1994-2015 by John E. Stone.
See the Tachyon distribution for redistribution and licensing information.
\index{rendering!Tachyon}
% Permission is granted to use Tachyon freely with VMD. -- JES 2006

\LISTITEM{Desmond and Maestro plugins by D. E. Shaw Research}
Copyright 2009, D. E. Shaw Research, LLC
All rights reserved.
\\
Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions are
met:
\\
  Redistributions of source code must retain the above copyright
  notice, this list of conditions, and the following disclaimer.
\\
  Redistributions in binary form must reproduce the above copyright
  notice, this list of conditions, and the following disclaimer in the
  documentation and/or other materials provided with the distribution.
\\
  Neither the name of D. E. Shaw Research, LLC nor the names of its
  contributors may be used to endorse or promote products derived from
  this software without specific prior written permission.
\\
THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
"AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
OWNER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
(INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.


\end{itemize}

