%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cr                                                                       
%cr            (C) Copyright 1995 The Board of Trustees of the            
%cr                        University of Illinois                         
%cr                         All Rights Reserved                           
%cr                                                                       
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RCS INFORMATION:
%
%       $RCSfile: ig.tex,v $
%       $Author: johns $        $Locker:  $                $State: Exp $
%       $Revision: 1.55 $      $Date: 2014/12/29 03:25:57 $
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DESCRIPTION:
%
% INSTALLATION GUIDE : MAIN LATEX File
%   Defines name, description, and sections for the installation
% guide.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



% LaTeX 2.09
%\documentstyle[11pt,html,graphic]{article}

% LaTeX 2E
\documentclass[11pt]{article}
\usepackage{html}
\usepackage{graphics}

% define margins, etc
\topmargin      0.1in
\oddsidemargin  0in
\evensidemargin 0in
\textheight     8.80in
\textwidth      6.50in
\marginparsep   0.25cm
\headheight     0in
\headsep        0in
%\footskip      0.5in
%\footheight    0in%

%
% the document itself
%


% define macros
\input{vmd_macros}

\newcommand{\DOCTITLE} {Installation Guide}

\newcommand{\DOCDESC} {
This document describes how to install one of the precompiled releases
of VMD and contains links to information on compilation of VMD from the
source code release.}

\begin{document}

\thispagestyle{empty}

\vspace*{0.3in}

\begin{center}
  \rule{6in}{0.04in}                            \\      \vspace{0.25in}
  {\Huge \VMD\ \DOCTITLE}                       \\      \vspace{0.25in}
  \rule{6in}{0.04in}                            \\      \vspace{0.25in}
  {\Large Version \VMDVER}                      \\      \vspace{0.20in}
  \VMDDATE                                      \\      \vspace{0.20in}
  \rule{6in}{0.04in}                            \\      \vspace{0.25in}

%  \htmladdimg{./vmd.gif}

  {\Large NIH Biomedical Technology Research Center for Macromolecular Modeling and Bioinformatics}  \\      \vspace{0.20in}
  {\large       \htmladdnormallinkfoot{Theoretical and Computational Biophysics Group}{http://www.ks.uiuc.edu/}  \\
                Beckman Institute for Advanced Science and Technology \\
                University of Illinois at Urbana-Champaign            \\
                405 N. Mathews                                        \\
                Urbana, IL  61801                                     \\
                \vspace{0.1in} \htmladdnormallink{{\tt http://www.ks.uiuc.edu/Research/vmd/}}{http://www.ks.uiuc.edu/Research/vmd/} \\
  }
  \vspace{0.5in}
  {\Large \bf Description}
  \vspace{0.1in}
\end{center}

{\DOCDESC}
\vspace{0.1in} \\
VMD development is supported by the National Institutes of Health
under grant numbers NIH~9P41GM104601 and 5R01GM098243-02.

\newpage
\input{vmd_ref}

\newpage
\section{Obtaining VMD Source and Binary Distributions}
The VMD source code and binary distributions can be obtained 
after registering at the VMD web page.  
Download the appropriate distribution file with your web browser.
Windows binary distributions are self extracting, so once the
distribution file is downloaded, proceed to the installation directions
below.

For source distributions and Unix binary distributions, uncompress and 
untar the file.  This will produce a subdirectory named {\tt vmd-\VMDVER}.  
Unless otherwise specified, all references to VMD code will be from this 
subdirectory, so {\tt cd} there.

\section {Installing a Pre-Compiled Version of VMD}
To install the pre-compiled Windows version of VMD,
simply run the self-extracting executable, and it will start the
VMD Windows installer program, which includes built-in help.
This process is automated and should be familiar to most
Windows users.  When installing VMD be sure that you have administrator
privileges.

To install the pre-compiled MacOS X bundle version of VMD, open the
VMD disk image and drag the VMD application into an appropriate 
directory.  Once the VMD application has been placed appropriately
it should be ready for immediate use as no other installation
steps are required.

To install the pre-compiled Unix version of VMD, then
only three steps remain to be done after you uncompress
and untar the distribution. 
\begin{itemize}
\item Edit the {\tt configure} script.  If necessary, 
change the following values:
{\tt
\begin{verbatim}
     $install_bin_dir
\end{verbatim}
}
\begin{verbatim}
	  This is the location of the startup script 'vmd'.  It should
	  be located in the path of users interested in running VMD.
\end{verbatim}
{\tt
\begin {verbatim}
     $install_library_dir
\end{verbatim}
}
\begin{verbatim}
	  This is the location of all other VMD files.  This includes
	  the binary and helper scripts.  It should not be in the path.
\end{verbatim}
\item Next generate the Makefile based on these configuration
variables.  This is done by running {\tt ./configure} .
\item After configuration is complete, {\tt cd} to the {\tt src}
directory and type {\tt make install}.  This will put the code in
the two directories listed above.  After this, you just type {\tt vmd}
to begin, provided that {\tt vmd} is in your path.
\end{itemize}


\input{ig_custom.tex}

\section{What to Do If It Doesn't Work}

  If you are running a VMD binary which has been built with a
native OpenGL implementation (i.e. not Mesa), you should make sure
that you have the vendor provided OpenGL runtime libraries and the
X server extensions correctly installed.  SGI systems normally have
the OpenGL runtime support installed on them.  Sun, HP, and IBM systems
often do not come with OpenGL support by default.  If you don't have the 
OpenGL runtime libraries for these systems, they can be downloaded for free
from the Sun, HP, and IBM web sites respectively.  Each of the vendor's OpenGL 
implementations generally include "install check" programs which verify
the correct installation and operation of the OpenGL libraries.

\begin{itemize}
\item Sun's OpenGL site is at {\tt http://www.sun.com/software/graphics/opengl/}
\item HP's OpenGL site is at {\tt http://www.hp.com/unixwork/products/opengl.html} 
\item IBM's OpenGL site is at {\tt http://www.austin.ibm.com/software/OpenGL/}
\end{itemize}
We suggest that you check that you are doing everything correctly,
and if it still doesn't work, report the problem by e-mail to
{\tt vmd@ks.uiuc.edu}. We'll try to help you.

 \section {Compiling Your Own Version of VMD}
   If for some reason you want to recompile VMD, then you will need to read 
 the rest of this document.  Most users will want to use the binary 
 distributions we provide since they have been thorougly tested prior to
 release.  It may be necessary for you to compile your own version of VMD
 in cases where we do not provide a binary for your platform, or
 when the provided binaries will not run correctly with a particular
 version of your operating system.  
 Full compilation instructions for VMD are found in the 
 online VMD Programmer's Guide:
   {\tt http://www.ks.uiuc.edu/Research/vmd/doxygen/}

\end{document}



