%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%cr                                                                       
%cr            (C) Copyright 1995 The Board of Trustees of the            
%cr                        University of Illinois                         
%cr                         All Rights Reserved                           
%cr                                                                       
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RCS INFORMATION:
%
%       $RCSfile: ig_custom.tex,v $
%       $Author: jordi $        $Locker:  $                $State: Exp $
%       $Revision: 1.20 $      $Date: 2003/04/09 21:51:52 $
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DESCRIPTION:
%
% INSTALLATION GUIDE : customizing VMD
%   1) .vmdrc file
%   2) .vmdsensors file
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\chapter{Customizing \VMD}
%\label{chapter:ig:custom}

\section{Customizing \VMD\ Startup}

The Unix version of \VMD\ reads in several data files (if they exist) 
when it starts up.
These files control the initial appearance and behavior of \VMD\ at
the start, and may be customized to suit each users particular tastes.
Default versions of these files are placed in the {\tt
INSTALLLIBDIR} directory (usually {\tt /usr/local/lib/vmd}). 
While each user may specify to use different versions of these files,
unless this is done the commands and values in the default files are
used.  In this way, an administrator may customize the default
behavior of \VMD\ for all users, while allowing each user the option
to change the default behavior however they choose.  This chapter
describes each of these data files.

Several configurable parameters may also be set in a number of ways, including by command-line options or by environment variables.  The order of precedence of these methods is as follows (highest precedence to lowest):
\begin{enumerate}
  \item Command-line options (see the Users Guide).
  \item Environment variable settings (see the Users Guide).
  \item Built-in defaults, as specified by compilation configurable parameters.
        These are used only if no other values are specified by the other 
        methods mentioned in this list.
\end{enumerate}

%\begin{centering}
%  \displayepsf[htb] (pictures/screen_params.eps scaled 800) {
%    \caption{Relationship between screen height, screen distance to origin,
%			and the viewer.}
%    \label{fig:ig:screen}
%  }
%\end{centering}

\section{The {\tt .vmdrc} and {\tt vmd.rc} files}

After initialization is complete, \VMD\ reads the {\em startup} file.  This file contains text commands
for \VMD\ to execute, just as if they had been entered at the \VMD\
text console command prompt.  The file can contain any number of
commands, including blank lines and comment lines (which begin with
the {\tt \#} character).  If an error is encountered while reading
this file, the command in error is skipped and processing of the file
continues.

The base filename for this startup file is {\tt .vmdrc} by default
on Unix systems and {\tt vmd.rc} on Windows;
this is determined by the configuration parameter {\tt STARTUPFILENAME}.
\VMD\ searches for this file in a number of locations,
and reads in the {\em first} version of the file it finds.  The order
of searching for the file is:
\begin{enumerate}
  \item {\tt ./STARTUPFILENAME}
  \item {\tt \$HOME/STARTUPFILENAME}
  \item {\tt INSTALLLIBDIR/STARTUPFILENAME}
\end{enumerate}
See the Users Guide for a list of all \VMD\ text commands.

\section{The {\tt .vmdsensors} file}

If \VMD\ is compiled with the VRPN option, it will look for files that specify how to access the external spatial
tracking devices. These files are read whenever \VMD\ is told to
initialize a specific external device.  The Tracker library will load 
the first file it finds in the following search order:
\begin{enumerate}
  \item {\tt \$HOME/.vmdsensors}
  \item The {\tt \$VMDSENSOR}  environment variable.
  \item {\tt INSTALLLIBDIR/.vmdsensors}
\end{enumerate}
%NOT TRUE RIGHT NOW: In the event of duplicated information, the device specifications found in 
%{\tt \$HOME/.vmdsensors} have precedence over those in {\tt \$VMDSENSOR}.

This last file ({\tt INSTALLLIBDIR/.vmdsensors}) contains extensive 
comments on how to configure the sensor description files
properly. If the VRPN option is omitted when compiling \VMD, 
this file is not used.
